Title here

### 2016-09-27 portfolio/text/ms.Rmd
Major in progress reworking of introduction

A fundamental challenge for individuals who rely on the extraction of natural
resources for their livelihood is income variability.
Fishers, loggers, farmers, miners, and others
all face uncertainty in income driven by a mix of
environmental, biological, and economic variability.
A key component underlying the sustainability
of these industries is identifying both
how regulatory actions and individual behaviour affect income variability
and by extension the economic stability and
sustainable use of natural resources.
For commercial fisheries, recent work has identified that
targeting a diversity of species can reduce income variability.
However, any diversity-variability relationship
is embedded in a regulatory framework that restricts individual ability
to unilaterally change fishing behaviour.
Here we identify three primary axes by which fishers in Alaska,
one of the world's most productive fishery regions, can
potentially diversify fishing activities and
investigate how individual actions along each diversity axis affect
both revenue and revenue variability.
We construct a novel statistical model to simultaneously
assess the value to individual fishers of choosing to
(1) own permits that allow targeting a diverse fish species assemblage,
(2) diversify by owning multiple permits, or
(3) catch a greater species diversity within the constraints of a given permit.
We show contrasting effects of diversity on
revenue and revenue variability along the three axes.
First, we show that fishing permits with
higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity).
Next, we show that fishing an additional permit
is associated with higher revenue
and lower variability (e.g.,
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits).
Finally, we show that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue the following year.
Thus, while there are clear benefits
to some axes of diversification, others are less effective.
Importantly, the most effective measure for
increasing revenue and decreasing risk ---
purchasing additional permits to participate in additional fisheries ---
was also the most costly.

At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies,
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

\bigskip
\noindent
Extra:

@@

Natural resources such as forests, land animals, and fishes
provide a primary source of income for millions of people worldwide [REFS].
However, environmental,
biological, and price variation
combined with regulatory changes
and human-caused environmental disasters
generate high levels of income variability and risk
for these individuals.
For example, drought and flooding are major sources of risk
to agricultural food security [@gregory2005],
And EVOS example

Increasing diversity has been proposed as a risk-buffering mechanism...
Give examples from other sectors

- Farming: crop rotation has been proposed as metric for buffering against short term
  variation and long term trends (like climate change)
- Logging: Maintaining a diverse mix of tree species, age structure, and space
  to buffer against things like wildfires and price variation.
- but in some situations, external forces (prices, demand) may discourage the
  adoption of these diverse strategies and
  help promote specialization [@bradshaw2004]

This effect has been identified for fisheries.
At community level, ... [@sethi2014]
At vessel level, ... [@kasperski2013] ...
However, these papers look at aggregates across individuals and largely ignore
how existing fisheries regulations limit the ability of individual fishers to
diversify. Key uncertainties remain about the fundamental relationship between
diversity, fishing revenue, and fishing revenue variability at the level of the
individual.
While the days of open, unregulated fisheries in most developed countries have
passed - fishers must own a license to participate in virtually all fisheries
- fishers retain several options for increasing diversity. For example, fishers
can do thing X, thing Y, or thing Z. But uncertainty remains about the value
of adopting different diversification strategies for improving revenue and
reducing revenue variability. Furthermore, theory and empirical studies both
suggest there can be substantial benefits to specialization (REFS?)

@@

Fisheries in Alaska are the highest valued fisheries in N. America
with landings valued at XX $ billion annually representing harvests from YY
species.  Alaska fisheries support a diverse fishing fleets ranging from small,
open, single operator vessels to multi-million dollar state of the art ships
with large crews. Similar diversity in fishing strategies can be found within
the Alaska fleet with different individuals adopting fishing strategies that
range from highly specialized (e.g. fishing only one species with one gear type
or in one area) to highly diverse (e.g. fishing multiple species with multiple
gear types in multiple areas). This broad range of fishing types provide an
unique opportunity for understanding the role diversity plays in determining
both revenue and risk for individual fishers.

We develop a novel statistical approach for understanding the effect of
diversity on both revenue and risk that builds on results from the vessel
[@kasperski2013], fleet [@sethi2012b], and community levels [@sethi2014].
First, we  summarize trends in diversification across fishing fleets in Alaska.
Next, we identified three main axes by which fishers may diversify: 1), 2), 3).
Finally, we constructed a model to simultaneously estimate the relationship
between each of these three axes of diversity and both annual revenue and risk
for individual fishers over nearly ZZ years.
Our hierarchical model makes use of
entire time series,
controls for effort and latent time effects,
and allows the effect of species diversity
to vary by fishing strategy (sets of similar permits).
It is the first model that jointly explores the effect of diversity on revenue
and risk at the individual level and the first considers the multiple axes of
diversity simultaneously.

>>>>

A fundamental challenge for individuals who rely on the extraction of natural
resources for their livelihood is income variability.
Fishers, loggers, farmers, and miners, among others,
face income uncertainty driven by a mix of
environmental, biological, and economic variability.
A key component underlying the sustainability
of these industries is identifying both
how regulatory actions and individual behaviour affect income variability
and by extension the economic stability and
sustainable use of natural resources.
For commercial fisheries, recent work has identified that
targeting a diversity of species can reduce income variability.
However, any diversity-variability relationship
is embedded in a regulatory framework that restricts individual ability
to unilaterally change fishing behaviour.

Here we identify three primary axes by which fishers in Alaska,
one of the world's most productive fishery regions, can
potentially diversify fishing activities and
investigate how individual actions along each diversity axis affect
both revenue and revenue variability.
We construct a novel statistical model to simultaneously
assess the value to individual fishers of choosing to
(1) own permits that allow targeting a diverse fish species assemblage,
(2) diversify by owning multiple permits, or
(3) catch a greater species diversity within the constraints of a given permit.

We show contrasting effects of diversity on
revenue and revenue variability along the three axes.
First, we show that fishing permits with
higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity).
Next, we show that fishing an additional permit
is associated with higher revenue
and lower variability (e.g.,
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits).
Finally, we show that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue the following year.

Thus, while there are clear benefits
to some axes of diversification, others are less effective.
Importantly, the most effective measure for
increasing revenue and decreasing risk ---
purchasing additional permits to participate in additional fisheries ---
was also the most costly.
At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies,
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

\bigskip
\noindent

Extra:

@@

Managing natural resources and the 
livelihoods of those that depend on their extraction is hard.
Environmental, biological, and economic variation
combine to generate high levels of income variability and risk.
Drought and flooding is a major source of 
risk to agricultural food security and farmers' incomes [@gregory2005].
Bark beetle outbreaks can cause devastation to 
the income of loggers.
Price crashes in insert mineral here 
are hard on miners' income variability ...

Individuals naturally develop strategies to reduce this income variation.
For example, farmers rotate their crops to buffer short-term 
climate variation and long-term environmental trends REF. 
Loggers maintain a diverse mix of tree species, age structure, 
geographical areas to buffer against 
disease outbreak, wildfires, and price variation REF. 
But, otherwise well-intentioned regulations
can limit the basket of actions
available to individuals to reduce income variability.
For example, ...
In other cases, external forces (prices, demand) may discourage the
adoption of these diverse strategies and
help promote specialization [@bradshaw2004].

Fisheries provide a prime example...
While the days of open, unregulated fisheries in most developed countries have
passed --- fishers must own a license to participate in virtually all fisheries
--- fishers retain several options for increasing diversity. 
For example, fishers can
fish multiple species as permitted by a license, 
own multiple licenses,
or fish in multiple regions.

But uncertainty remains about the value
of adopting various diversification strategies for improving revenue and
reducing revenue variability. 
Furthermore, theory and empirical studies both
suggest there can be substantial benefits to specialization (REFS).

However, these papers look at aggregates across individuals and largely ignore
how existing fisheries regulations limit the ability of individual fishers to
diversify. 
Key uncertainties remain about the fundamental relationship between
diversity, fishing revenue, and fishing revenue variability at the level of the
individual.

Diversity in the fishing strategies of individuals
ranges from highly specialized (e.g. fishing only one species with one gear type
or in one area) to highly diverse (e.g. fishing multiple species with multiple
gear types in multiple areas).

Fisheries in Alaska provide a unique opportunity 
to understand the role of individuals' strategies 
in determining income and income variability
for those whose livelihood depends on the extraction of natural resources.
Fisheries in Alaska are the highest-valued fisheries in North America
with landings valued at XX $ billion annually representing harvests from YY
species.
The industry supports a range of fishing fleets ranging from small,
open, single-operator vessels to multi-million-dollar state-of-the-art ships
with large crews.

However, the days of open, unregulated fisheries in most developed countries have
passed --- fishers must own a license to participate in virtually all fisheries.
And while such regulations have proven 
helpful in reducing or eliminating overfishing,
they do limit the available strategies to reduce income variability.

@@

We develop a novel statistical approach for understanding the effect of
diversity on both revenue and risk that builds on results from the vessel
[@kasperski2013], fleet [@sethi2012b], and community levels [@sethi2014].
First, we  summarize trends in diversification across fishing fleets in Alaska.
Next, we identify three main axes by which individual fishers may diversify
the species they target:
(1) owning permits that allow targeting a diversity of fish species,
(2) purchasing additional permits,
or (3) catching a greater species diversity within the constraints of a given permit.

Finally, we construct a model to simultaneously estimate the relationship
between each of these three axes of diversity and both annual revenue and risk
for individual fishers over nearly ZZ years.
Our hierarchical model makes use of
entire time series,
controls for effort and latent time effects,
and allows the effect of species diversity
to vary by fishing strategy (sets of similar permits).
It is the first model that jointly explores the effect of diversity on revenue
and risk at the individual level and the first considers the multiple axes of
diversity simultaneously.
